Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-10-02 arxiv/proof_examples.tex
revise

@@

		\hat{B}(A) \define  (A_1^2 (X^{(1)})^{\top}X^{(1)} + A_2^2 (X^{(2)})^{\top}X^{(2)})^{-1} (A_1 (X^{(1))})^{\top}Y^{(1)} + A_2 (X^{(2)})^{\top}Y^{(2)}). \label{eq_Bhat_2task} %\\

@@

 Now inserting $\hat B$ back into \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $A:=A_1/A_2$ as follows 

@@

		 g(A) \define & \left\| X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\hat \Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\hat \Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)+ \left(X^{(2)}\hat \Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} + A X^{(2)}\hat \Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 

@@

	$$\hat \Sigma:= A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$

@@

&= \left\| X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\hat \Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon_1} +\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\hat \Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \\
&+  A^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\hat \Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)}}  +A^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  (X^{(2)})^\top X^{(2)}\hat \Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)}  }\\
		&= \left\| X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\hat \Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\hat \Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\hat \Sigma^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).

@@

&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

 &\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \hat \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] \\
 &=\tr\left[  (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} \right] \\
 & = \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right] .

@@

 $$(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \hat \Sigma^{-1}(X^{(1)})^\top X^{(1)}.$$

@@

\begin{proof}
Corresponding to \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}. \label{op_uplow}
\end{align}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we used \eqref{choiceofpara0}. Hence using Claim \ref{claim_largedev1}, \eqref{eq_h1} and \eqref{eq_hA}, we obtain that with high probability,
\begin{align}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1) \label{gA>g1}
\end{align}
under \eqref{choiceofpara0}.
This gives that
\be\label{rough A-1}
|\hat A-1|\le \delta
\ee
for any small constant $\delta>0$.

@@

%Using Lemma \ref{largedeviation} again, we can simplify \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as

@@

%	Furthermore, the validation loss in \eqref{approxvalid} reduces to

@@

%	\item Second, with \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get \eqref{simple1}.

@@

h(A):= [(1-A^{-1})^2\kappa^2 + (1+A^{-2})d^2/2] \cdot p^{-1} \tr\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right).

@@

$$\tr\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right).$$

@@

h'(A) \ge \left[2(1-A^{-1})\frac{\kappa^2}{A^2} - \frac{d^2}{A^3}\right] \cdot p^{-1} \tr\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right) \ge 0,

@@

%Using the estimate Fact \eqref{lem_minv}, we get that $\cal A$ is lower bounded as

@@

$$\tr\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} +  [(X^{(1)})^\top X^{(1)}]^{-1}\right).$$

@@

	h'(A)\le [-2(1-A)\kappa^2 + Ad^2] \cdot p^{-1} \tr\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)\le 0,

@@

%In sum, we have shown that $w_2\le w\le w_1$. Together with
%$$\max\{|w_1 -1|, |w_2 -1|\} =\OO\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right),$$
%we conclude \eqref{hatw_add1}.

@@

Now we are ready to prove \eqref{hatw_add1}. First using \eqref{op_uplow} and \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le A\le 1+\delta$,
$$h'(A)\ge \frac{2(A-1)\kappa^2 - d^2}{A^3} \cdot p^{-1} \tr\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right) \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+e_2}\le A\le 1+\delta$,
\begin{align*}
h(A)-h(1)\ge h(A)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^A h'(x)\dd x \gtrsim p\kappa^2 \cdot |A-1|^2  \gtrsim p\frac{d^2}{\kappa^2} + p^{-1/2+2e_2}\cdot p\kappa^2.
\end{align*}
Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
$$ h(A)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under \eqref{choiceofpara0}, which implies the inequality as in \eqref{gA>g1}. Second using \eqref{op_uplow} and \eqref{h'(A)2}, we obtain that for $1-\delta \le A \le 1-d^2/\kappa^2$,
$$- h'(A)\ge  [2(1-A)\kappa^2 - Ad^2] \cdot p^{-1} \tr\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)  \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-\frac12c_\infty}\le A\le 1-\delta$, then 
$$ h(A)-h(1) \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under \eqref{choiceofpara0}, which implies the inequality as in \eqref{gA>g1}. In sum, we obtain that the minimizer $\hat A$ must satisfy \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with \eqref{hatw_add1}, we conclude \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Calculating the prediction loss.} Now we calculate the predication loss using the optimal $\hat B(\hat A) $ in \eqref{eq_Bhat_2task}. ......

>>>>

\iffalse
%Our key insight is a bias-variance decomposition of the expected prediction loss of $\hat{\beta}_2^{\MTL} = \hat{B} W_2$.
%Using the local optimality condition of $B$, we obtain that
%\begin{align*}
%	 \hat{B}(W_1, W_2) &= (W_1^2 X_1^{\top}X_1 + W_2^2 X_2^{\top}X_2)^{-1} (W_1 X_1^{\top}Y_1 + W_2 X_2^{\top}Y_2)\\
%	&= \frac{1}{W_2} \left( \frac{W_1^2}{W_2^2}  X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1} \left(\frac{W_1}{W_2} X_1^{\top}Y_1 + X_2^{\top}Y_2\right) \\
%	&= \frac{1}{W_2}\left[\beta_2 + \left(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2\right)^{-1}\bigbrace{X_1^{\top}X_1\left(\frac{W_1}{W_2}\beta_1 - \frac{W_1^2}{W_2^2} \beta_2\right) + \left(\frac{W_1}{W_2} X_1^{\top}\varepsilon_1 + X_2^{\top}\varepsilon_2\right)}\right].
%\end{align*}
Recall that $\hat{\beta}_i^{\MTL} = \hat{B} W_i$. Therefore, we have
\begin{align}
	\exarg{\epsilon_1, \epsilon_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}
	=&~ \frac{W_1^2}{W_2^2} \bignorm{\Sigma_2^{1/2}(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} X_1^{\top}X_1 (\beta_1 - \frac{W_1}{W_2} \beta_2)}^2 \label{eq_bias_2task} \\
			&+~  \sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} }. \label{eq_var_2task}
\end{align}
equation \eqref{eq_bias_2task} is the bias of $\hat{\beta}_t^{\MTL}$ and
equation \eqref{eq_var_2task} is the variance of $\hat{\beta}_t^{\MTL}$.
%minus the variance of $\hat{\beta}_t^{\STL}$, which is always negative.
Comparing the above to single-task learning, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

% \todo{derive the test loss} Expectation was given in equation \eqref{eq_bias_2task} and equation \eqref{eq_var_2task}.
 \fi

@@

		\hat{B} \define  (A_1^2 (X^{(1)})^{\top}X^{(1)} + A_2^2 (X^{(2)})^{\top}X^{(2)})^{-1} (A_1 (X^{(1))})^{\top}Y^{(1)} + A_2 (X^{(2)})^{\top}Y^{(2)}). \label{eq_Bhat_2task} %\\

@@

 Now inserting $\hat B$ back into equation \eqref{eq_mtl_2task_cov}, we obtain an objective that only depends on $A:=A_1/A_2$ as follows 

@@

		 g(A) \define & \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)+ \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)\epsilon^{(1)} + A X^{(1)}\Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)} \right\|^2 \nonumber\\
		 + & \left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)+ \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)\epsilon^{(2)} + A X^{(2)}\Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)} \right\|^2, 

@@

	$$\Sigma\equiv \Sigma(A):= A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$

@@

&= \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
& + \exarg{\epsilon^{(1)}}{(\epsilon^{(1)})^\top \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\epsilon_1} +\exarg{ \epsilon^{(2)}}{(\epsilon^{(2)})^\top \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2\epsilon^{(2)}}   \\
&+  A^2 \exarg{ \epsilon^{(2)}}{ (\epsilon^{(2)} )^\top X^{(2)} \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\Sigma^{-1} (X^{(2)})^\top \epsilon^{(2)}}  +A^2 \exarg{ \epsilon^{(1)}} { (\epsilon^{(1)})^\top X^{(1)} \Sigma^{-1}  (X^{(2)})^\top X^{(2)}\Sigma^{-1} (X^{(1)})^\top \epsilon^{(1)}  }\\
		&= \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (A\beta_2-\beta_1)\right\|^2 +\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (A\beta_1-A^2\beta_2)\right\|^2  \\
		&+ \sigma^2\tr\left[ \left(A^2 X^{(1)}\Sigma^{-1} (X^{(1)})^\top-\id_{n_1}\right)^2\right] +\sigma^2\tr\left[ \left(X^{(2)}\Sigma^{-1} (X^{(2)})^\top-\id_{n_2}\right)^2 \right]\\
&+ 2 A^2 \sigma^2 \tr\left[ \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\Sigma^{-1} (X^{(2)})^\top X^{(2)} \right]  \nonumber\\
		& = \left\| X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)} (\beta_1-A\beta_2)\right\|^2 +A^2\left\| X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)} (\beta_1-A\beta_2)\right\|^2 +\sigma^2(n_1+n_2-p).

@@

&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right] \nonumber\\
&+ A^2[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1}\tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] +\sigma^2(n_1+n_2-p) \nonumber\\
&=[(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right]+\sigma^2(n_1+n_2-p), \label{EgA}

@@

 &\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \Sigma^{-1} (X^{(1)})^\top X^{(1)}\right] \\
 &=\tr\left[  (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1} (X^{(2)})^\top X^{(2)}\right]+ A^2 \tr\left[  (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} \right] \\
 & = \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)}\right] .

@@

 $$(X^{(1)})^\top X^{(1)} \Sigma^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} = (X^{(2)})^\top X^{(2)} \Sigma^{-1}(X^{(1)})^\top X^{(1)}.$$

@@

\begin{proof}
Corresponding to equation \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}. \label{op_uplow}
\end{align}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(1)^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we used equation \eqref{choiceofpara0}. Hence using Claim \ref{claim_largedev1} and equations \eqref{eq_h1}-\eqref{eq_hA}, we obtain that with high probability,
\begin{align}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1) \label{gA>g1}
\end{align}
under condition \eqref{choiceofpara0}.
This gives that
%\be\label{rough A-1}
$|\hat A-1|\le \delta$ for any small constant $\delta>0$.

@@

%Using Lemma \ref{largedeviation} again, we can simplify equation \eqref{revise_eq_val_mtl} as $\val(v)= N_2h(v) \cdot  \left( 1+\OO(p^{-1/2+\e})\right)$, where the function $ h$ is defined as

@@

%	Furthermore, the validation loss in equation \eqref{approxvalid} reduces to

@@

%	\item Second, with equation \eqref{boundv-w} we show that $\hat v$ is close to $\hat w$. Then we plug $\hat{v}$ into $\te(\hat{\beta}_2^{\MTL})$ to get equation \eqref{simple1}.

@@

h(A):= [(1-A^{-1})^2\kappa^2 + (1+A^{-2})d^2/2] \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right].

@@

$$\tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]$$

@@

h'(A) \ge \left[2(1-A^{-1})\frac{\kappa^2}{A^2} - \frac{d^2}{A^3}\right] \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \ge 0,

@@

%Using the estimate Fact equation \eqref{lem_minv}, we get that $\cal A$ is lower bounded as

@@

$$\tr\left[\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} +  [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right].$$

@@

	h'(A)\le [-2(1-A)\kappa^2 + Ad^2] \cdot p^{-1} \tr\left[\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]\le 0,

@@

%In sum, we have shown that $w_2\le w\le w_1$. Together with
%$$\max\{|w_1 -1|, |w_2 -1|\} =\OO\left(\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2}\right),$$
%we conclude equation \eqref{hatw_add1}.

@@

Now we are ready to prove equation \eqref{hatw_add1}. First using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le A\le 1+\delta$,
$$h'(A)\ge \frac{2(A-1)\kappa^2 - d^2}{A^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+e_2}\le A\le 1+\delta$,
\begin{align*}
h(A)-h(1)\ge h(A)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^A h'(x)\dd x \gtrsim p\kappa^2 \cdot |A-1|^2  \gtrsim p\frac{d^2}{\kappa^2} + p^{-1/2+2e_2}\cdot p\kappa^2.
\end{align*}
Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
$$ h(A)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under equation \eqref{choiceofpara0}, which implies the inequality \eqref{gA>g1}. Second using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le A \le 1-d^2/\kappa^2$,
$$- h'(A)\ge  [2(1-A)\kappa^2 - Ad^2] \cdot p^{-1} \tr\left[\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-\frac12c_\infty}\le A\le 1-\delta$, then 
$$ h(A)-h(1) \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under equation \eqref{choiceofpara0}, which implies the inequality \eqref{gA>g1}. In sum, we obtain that the minimizer $\hat A$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}
